Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2015-05-24 structured.tex
2015.05.17 (Alex) Minor re-wording for the text on the 3DO algorithm. Added connection sentence for discussion on structured approaches between introductory paragraph and main body of the section. Some additional re-wording and figure re-scaling in the (structured) discussion text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{3D Overlay (3DO)}~\cite{AOS2014} targets mobile networks.
As the name suggests, the logical space of the network is 
a $3$--$d$ space and the used ``transient'' (continually changing) 
logical \emph{ID}s (\emph{LID}s) consist of $3$ axes, $x$, $y$ and $z$ 
in the form of a tuple;
each such axis value is represented by an $M$-bit identifier.
%%
A newlly-arriving peer broadcasts a join request using
an TTL-based expanding ring search. 
This helps locate peers that are physically adjacent.
Responding peers send in 
their \emph{LID}s along with a list of their neighbors 
and their corresponding distances from those neightbors. 
%%
Once the joining peer receives the above responses,
it stores requisite information in its own routing table 
and so, it generates an undirected connected
graph of itself with both its $1$- and $2$-hop neighbours. 
%%
Starting off this graph and with the help of hop 
distances obtained from the underlying MANET,
the incoming peer builds a minimum spanning tree (MST)
with itself as the root node; 
in this tree, the nodes adjacent to the root become the
direct neighbors of the peer.
%%
In addition, the entries of the initial routing table 
that do not correspond to above direct neighbors are
eliminated.

The procedure above ascertains that close peers on the
overlay are also adjacent at the physical layer. 
%%
At this stage, a peer can generate its \emph{LID}
that automatically places it in the $3$-$d$ network space.
Updates in the peer's connectivity infromation 
reflecting continual changes in the mobile network,
trigger updates on the composition of \emph{LID} itself.

%%AD work on what the following says.. 
%%VM I wrote the following paragraph as 
\emph{3DO} stores objects in the same $3$-$d$ space. For each
file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
is hashed out of the file's unique ID~(e.g., filename).
The node whose \emph{LID} is closer to $k$ stores the
file itself.
%%
To retrieve the file, the requesting node calculates $f$'s $k$
and issues a query message. The message is a associated with a TTL
value. Each peer forwards that query to the node,
in its routing table, that is closest to $k$ up until the peer
that stores the file is reached or the query's lifetime expires.

\emph{3DO} focuses on
the creation of a $3$-$d$ space that yields improved load balancing 
and uses fine--grained information to ensure
that each hop on the overlay during object location, 
is also closer physically to the sought object.
%%
In contrast, the use of TTL in query messages
renders the network not--a--pure DHT as it abolishes 
the zero false--negative queries that traditional DHTs maintain.
\cite{AOS2014} also concedes that the approach
does not rapidly adapt to network changes and so
its evaluation was solely based on low mobility scenarios. 
For the above reasons, \emph{3DO} exhibits the following behavior:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% 
low &
% 
low &
% 
medium
\end{tabular}
}
\end{center}

@@

\emph{SAT-Match}~\cite{RGJZ2004} defines an iterative process that 
adaptively changes the overlay network to ease the topology mismatch problem. 
It consists of a \emph{probing phase} in which a peer queries nearby nodes 
for distances and a \emph{jumping phase} in which the peer decides 
to change its neighborhood in order to reduce average neighbor distances. 
This process is performed iteratively
until no additional optimization is required; 
yet, the methods achieves optimization within a
sufficiently large scope.
%% VM This is used in other papers. Maybe we can put somewhere in the
% definition of Topology Mismatch Problem in the intro/background.
The notion of \emph{stretch} is used as the means for quantifying 
the \emph{topology match degree} of the constructed overlay;
stretch is defined as: 
$S = \bar{L_l}/\bar{L_p}$ where $\bar{L_l}$ is the average logical link
latency and $\bar{L_p}$ is the average physical link latency.
%%AD what are the indexes: l and p -- p for peer and l for new peer???
%% VM: no l for logical and p for physical
%%AD rephrased what is below as above... 
% 
% The paper defines \emph{stretch}
% $S = \bar{L_l}/\bar{L_p}$, where $\bar{L_l}$ is the average logical link
% latency while the $\bar{L_p}$ is the average physical link latency, and uses it
% a way to quantify the \emph{topology match degree} of the constructed overlay.

% The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
% using the standard mechanism of the algorithm at hand. 
When a peer joins the \emph{DHT}, it starts 
its \emph{probing phase} by contacting the neighborhood with 
messages having small \emph{TTL=k}~\cite{jiang_lightflood_2008}. 
Recipients return information to the source and forward the message 
if not expired. 
After discovering its \emph{TTL}-$k$ neighborhood, 
the source measures \emph{RTT}-times, sorts them
and picks the two nodes with the smallest \emph{RTT}s.

@@

So far in this section, we have discussed research efforts 
trying to address the issue of inefficient network resource utilization 
manifested by network--topology mismatches in structured \p-systems.
Based on the salient feature of outlined techniques
\cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
using the following aspects:
\begin{enumerate}[\itshape i\upshape)]
  \item geographic-layout,
  \item proximity routing, and 
  \item proximity neighbor selection,
\end{enumerate}

%%VM Added the following to mention the hybrid/combo methodologies
Similarly to unstructured approaches, most of the methods
reviewed in this section entail features from multiple of the above categories. 
For example, \cite{KLKP2008}~uses both proximity routing 
and proximity neighbor selection to achieve topology adaptation,
while \emph{LAPTOP} combines geographic--layout with
replication strategies to refine its results.
%%%
The heuristic nature of the suggested solutions opens them up 
to trade-offs which, in most cases, are 
highly affected by the application domain and the implementation itself.

Table~\ref{structured:table} offers a quick overview of how the surveyed 
approaches behave.
The topics of topology adaptation, landmarking and caching/replication
are not further elaborated as they have been discussed 
in Section~\ref{section:unstructured}.
In a similar fashion to Table~\ref{unstructured:table}, 
we outline key properties of each surveyed approach 
that helps diffferentiate techniques from competitors.
We also highlight pros and cons for all proposals discussed.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.36]{img/pdf/geographic-layout.pdf}
\caption{Successive nodes in the overlay circle can be chosen to be from the
same or near autonomous systems (\emph{AS}).}
\label{figure:geographic-layout}
\end{figure}
%%
A number of the reviewed algorithms including 
\emph{Global Soft-State},
\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
order to reduce the network's average delay.
Landmark servers and \emph{RTT} measurements are two popular methods, which
can also be used in conjunction, to discover peers in physical proximity.
However, as already discussed, these
methods do not always yield reliable estimates for node positions over the
Internet. 
Landmark servers are not self-organizing and present maintenance
overhead. Moreover, to serve a \p--network with potentially millions of peers, multiple
such servers are required to be distributed over the world, which is a difficult
proposition per se.
\emph{RTT}s can measure the delay between peers, but their use constitutes a
greedy method that can result in sub-optimal overlay topologies;
the latter is especially true if nearby 
nodes happen to be connected through low bandwidth connections.

On one hand, approaches based on \emph{geographic layout} improve the query
efficiency of the system in general and are especially effective when applied
to CAN-like implementations. On the other hand, they inherently undermine the
uniform distribution of peer identifiers. This is an important downside as it
creates hot-spots that hurt system scalability. Moreover, as nearby nodes
are more likely to suffer collective failures~\cite{HY2007}, it adversely affects
system availability.
%% VM reworded the above paragraph, please check it out.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-routing.pdf}
\caption{Proximity routing takes proximity information into account during the
routing process.}
\label{figure:proximity-routing}
\end{figure}
%%
We reviewed an array of methodologies that strictly adhere to 
the tenets of this category such as \emph{Mithos} and
\emph{PChord} as well as others that function along with 
proximity neighbor selection such as \emph{DynaMO}.
%%
When \emph{PR}--based algorithms route messages, they try to balance between
choosing the node that will further progress the routing towards the
destination and picking the closest entry in the routing table
in terms of network proximity. 
%%AD just above in this section you have not clarified the way 
%%	routing tables get populated - u have to clarify this point 
%%	in order for the contrast here to be valid..
%%
%% VM Is the above now better with the rewording of the section's opening paragraph?
In the degenerate case, a system could use a greedy approach to select
only the lowest latency candidate peers during forwarding. That would to
longer, sub--optimal, paths to the destinations; thus imposing
extra stress to network resources. PR is also less
effective than geographical layout when applied to CAN--like implementations.
%%AD re-work carefully this segment...
%% VM how is it now?

@@

% \emph{Proximity Neighbor Selection (PNS)} 
Algorithms in this category use proximity data during routing table construction
and maintenance along the lines of Figure~\ref{figure:proximity-neighbour-selection}.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/proximity-neighbor-selection.pdf}
\caption{Proximity neighbor selection dictates that proximity is taken into
account during routing table maintenance.}
\label{figure:proximity-neighbour-selection}
\end{figure}
%%
The proximity information used is different from that of 
landmark-based systems described in Section~\ref{section:geographic_layout};
here, \emph{RTT} values between nodes (i.e., \emph{SAT-Match}) or 
\emph{IP} address prefixes (i.e., \emph{DynaMO}) % or \emph{Cone}) %%VM removing Cone from paper
are predominantly used to detect proximity. 
\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger
than $24$ bits belong to a single geographical location. 
However, using a smaller number of bits creates less precise results 
and a larger number of bits may increase the burden in the network
and simultaneously reduce the possible number of neighbors. 
%%
Moreover, \cite{HLYW2005} argue that prefix--based approaches can
destroy the uniform distribution in the key space and do not work in
one-dimensional key space where the mapping is restricted to 
operate at the overlay.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{3D Overlay (3DO)}~\cite{AOS2014} seeks to address the 
network mismatch problem  for peers operating in mobile networks.
As its name suggests, the network logical space is 
a $3$--$d$ space and the used ``transient'' (continually changing) 
logical \emph{ID}s (\emph{LID}s) consist of $3$ axes, $x$, $y$ and $z$ 
in the form of a tuple;
each such axis value is represented by an $M$-bit identifier.
%%
A newlly-arriving peer broadcasts a join request using
an \emph{TTL}-based expanding ring search helping locate 
peers that are physically adjacent.
Responding peers send in 
their \emph{LID}s along with a list of their neighbors 
and their corresponding distances from those neightbors. 
%%
Once a joining peer receives the above responses,
it stores requisite information in its own routing table 
and so, it generates an undirected connected
graph of itself with both its $1$- and $2$-hop neighbours. 
%%
Starting off this graph and with the help of hop 
distances obtained from the underlying MANET,
the incoming peer builds a minimum spanning tree (MST)
with itself as the root node; 
in this MST, the nodes adjacent to the root become the
direct neighbors of the peer.
%%
Moreover, the entries of the initial routing table 
that do not correspond to above direct neighbors are
eliminated.

The above procedure ascertains that close peers on the
overlay are also adjacent at the physical layer. 
%%
At this stage, a peer can generate its \emph{LID}
that automatically places it in the $3$-$d$ network space.
Updates in the peer's connectivity infromation 
reflecting continual changes in the mobile network,
trigger updates on the composition of \emph{LID} itself.

\emph{3DO} stores objects in the same $3$-$d$ space. For each
file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
is hashed out of the file's unique \emph{ID}~(e.g., filename).
The node whose \emph{LID} is closer to $k$ stores the
file itself.
%%
To retrieve the file, the requesting node calculates $f$'s $k$
and issues a query message. 
The message is a associated with a \emph{TTL} value. 
Each peer forwards that query to the node,
in its routing table, that is closest to $k$ until either the peer
that stores the file is located or the query's lifetime expires.

\emph{3DO} focuses on
the creation of a $3$-$d$ space whose stated objective is to
yield improved load balancing 
and exploit fine--grained information to ensure
that each hop on the overlay during object location, 
is also closer physically to the sought object.
%%
However, the use of \emph{TTL} in query messages
renders the network not--a--pure DHT as it abolishes 
the zero false--negative queries that traditional DHTs maintain.
\cite{AOS2014} also concedes that the approach
does not rapidly adapt to network changes and so
its evaluation was solely based on low mobility scenarios. 
For the above reasons, \emph{3DO} exhibits the following behavior 
in conjunction with our three stated criteria:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% 
low &
% 
low &
% 
medium
\end{tabular}
}
\end{center}

@@

\emph{SAT-Match}~\cite{RGJZ2004} defines an iterative process that 
adaptively changes the overlay network to ease the topology mismatch problem. 
It consists of a \emph{probing phase} in which a peer queries nearby nodes 
for distances and a \emph{jumping phase} in which the peer decides 
to change its neighborhood in order to reduce average neighbor distances. 
This process is performed iteratively
until no additional optimization is required; 
yet, the methods achieves optimization within a
sufficiently large scope.
%% VM This is used in other papers. Maybe we can put somewhere in the
% definition of Topology Mismatch Problem in the intro/background.
The notion of \emph{stretch} is used as the means for quantifying 
the \emph{topology match degree} of the constructed overlay;
stretch is defined as: 
$S = \bar{L_l}/\bar{L_p}$ where $\bar{L_l}$ is the average logical link
latency and $\bar{L_p}$ is the average physical link latency.
% 
% The \emph{probing phase} starts when the new-comer joins the \emph{DHT} 
% using the standard mechanism of the algorithm at hand. 
When a peer joins the \emph{DHT}, it starts 
its \emph{probing phase} by contacting the neighborhood with 
messages having small \emph{TTL=k}~\cite{jiang_lightflood_2008}. 
Recipients return information to the source and forward the message 
if not expired. 
After discovering its \emph{TTL}-$k$ neighborhood, 
the source measures \emph{RTT}-times, sorts them
and picks the two nodes with the smallest \emph{RTT}s.

@@

In this section, we have discussed research efforts 
trying to address the issue of inefficient network resource utilization 
manifested by network--topology mismatches in \emph{structured} \p-systems.
Based on the salient feature of outlined techniques
\cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
based on the following aspects:
\begin{enumerate}[\itshape i\upshape)]
  \item geographic-layout,
  \item proximity routing, and 
  \item proximity neighbor selection,
\end{enumerate}
%%
%%COM Added the following to mention the hybrid/combo methodologies
Similarly to unstructured approaches, most of the methods
reviewed in this section entail features from multiple of the above categories. 
For example, \cite{KLKP2008}~uses both proximity routing 
and proximity neighbor selection to achieve topology adaptation,
while \emph{LAPTOP} combines geographic--layout with
replication strategies to refine its results.
%%%
The heuristic nature of the suggested solutions opens them up 
to trade-offs which, in most cases, are 
highly affected by the application domain and the implementation itself.

Table~\ref{structured:table} offers a comparative view 
on how our surveyed approaches behave.
The topics of topology adaptation, landmarking and caching/replication
are not further elaborated as they have been discussed 
in Section~\ref{section:unstructured}.
In a similar fashion to Table~\ref{unstructured:table}, 
we outline key properties of each surveyed approach 
that helps diffferentiate techniques from competitors.
We also highlight pros and cons for all proposals discussed.

%%AD Connection???  to the groups???
In what follows, we place the surveyed structured \p2p-methods in 
the three categories mentioned above and which are formulated based 
on utilized proximity and geographic aspects.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.44]{img/pdf/geographic-layout.pdf}
\caption{Successive nodes in the overlay circle can be chosen to be from the
same or near autonomous systems (\emph{AS}).}
\label{figure:geographic-layout}
\end{figure}
%%
A number of the reviewed algorithms including 
\emph{Global Soft-State},
\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
order to reduce the network's average delay.
Landmark servers and \emph{RTT} measurements are two popular methods, which
can also be used in conjunction, to discover peers in physical proximity.
However, as already discussed, these
methods do not always yield reliable estimates for node positions over the
Internet. 
Landmark servers are not self-organizing and present maintenance
overhead. Moreover, to serve a structured \p--network with potentially 
tens of thousands or more 
% millions 
peers, multiple such servers are required to be distributed over the world, 
which is a difficult proposition per se.
\emph{RTT}s can measure the delay between peers, but their use constitutes a
greedy method that can result in sub-optimal overlay topologies;
the latter is especially true if nearby 
nodes happen to be connected through low bandwidth connections.

On one hand, approaches based on \emph{geographic layout} improve the query
efficiency of the system in general and are especially effective when applied
to \emph{CAN}-like implementations. On the other hand, they inherently undermine the
uniform distribution of peer identifiers. This is an important downside as it
creates hot-spots that hurt system scalability. 
Moreover, it adversely affects system availability as nearby nodes
are more likely to suffer collective failures~\cite{HY2007}.

@@

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.48]{img/pdf/proximity-routing.pdf}
\caption{Proximity routing takes proximity information into account during the
routing process.}
\label{figure:proximity-routing}
\end{figure}
%%
We reviewed an array of methodologies that strictly adhere to 
the tenets of this category such as \emph{Mithos} and
\emph{PChord} as well as others that function along with 
proximity neighbor selection such as \emph{DynaMO}.
%%
When \emph{PR}--based algorithms route messages, they try to balance between
choosing the node that will further progress the routing towards the
destination and picking the closest entry in the routing table
in terms of network proximity. 
%%AD just above in this section you have not clarified the way 
%%	routing tables get populated - u have to clarify this point 
%%	in order for the contrast here to be valid..
%%
%% VM Is the above now better with the rewording of the section's opening paragraph?
In the degenerate case, a system could use a greedy approach to select
only the lowest latency candidate peers during forwarding. That would to
longer, sub--optimal, paths to the destinations; thus imposing
extra stress to network resources. PR is also less
effective than geographical layout when applied to CAN--like implementations.
%%AD re-work carefully this segment...
%% VM how is it now?

@@

% \emph{Proximity Neighbor Selection (PNS)} 
Algorithms in this category use proximity data during routing table construction
and maintenance along the lines of Figure~\ref{figure:proximity-neighbour-selection}.
%%
\begin{figure}[hbft]
\centering
  \includegraphics[scale=0.48]{img/pdf/proximity-neighbor-selection.pdf}
\caption{Proximity neighbor selection dictates that proximity is taken into
account during routing table maintenance.}
\label{figure:proximity-neighbour-selection}
\end{figure}
%%
The proximity information used is different from that of 
landmark-based systems described in Section~\ref{section:geographic_layout};
here, \emph{RTT} values between nodes (i.e., \emph{SAT-Match}) or 
\emph{IP} address prefixes (i.e., \emph{DynaMO}) % or \emph{Cone}) %%VM removing Cone from paper
are predominantly used to detect proximity. 
\cite{freedman_iploc_2005} states that $97\%$ of prefixes larger
than $24$ bits belong to a single geographical location. 
However, using a smaller number of bits creates less precise results 
and a larger number of bits may increase the burden in the network
and simultaneously reduce the possible number of neighbors. 
%%
Moreover, \cite{HLYW2005} argue that prefix--based approaches can
destroy the uniform distribution in the key space and do not work in
one-dimensional key space where the mapping is restricted to 
operate at the overlay.
